---
title: "Assignment 1"
output: pdf_document
date: "2023-01-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This is a data set holds all of the matches that world cup competitors participated in, their likelihood to win/lose/tie said match, and also specifies how likely each team will finish first or second in their group and eventually advance to the knockout stage. 

```{r wc_forecasts}
wc_forecasts <- read.csv("~/Masters/DATA607/wc_forecasts.csv")
summary(wc_forecasts)
```

## Subset of Data

I'm curious about South Korea in particular, so I'd like to see their journey throughout the world cup this past year, specifically regarding whether they'll make the quarters/semis/finals.

```{r subset}
library(dplyr)
relevant_data <- subset(select(wc_forecasts, c('forecast_timestamp', 'team', 'group', 'spi', 'make_round_of_16', 'make_quarters', 'make_semis', 'make_final', 'win_league')), team=="South Korea")
colnames(relevant_data)[colnames(relevant_data) == "spi"] = "soccer_power_index"
relevant_data
```

## Conclusion
We can see that South Korea was predicted to be somewhat successful and maybe get into quarterfinals from the first two games, but never really had a chance to win the league or even make the finals. It does look like they were unlikely to win their 11/28 game though, and them moving forward increased their chance of winning substantially, however they were knocked out in the next game and that ended their chance to win all together.