---
title: "Week 7 Assignment"
author: "Alice Ding"
date: "2023-03-07"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(rvest)
library(jsonlite)
library(xml2)
library(tibble)
```

## Overview

I've picked 3 fiction books, each written by people of color and telling stories from their own personal lens. These brings different cultures through the authors' own personal experiences, but each story is inherently American in nature as well. 

I've put the following information in `html`, `xml`, and `json` formats:

- Title
- Author(s)
- Publish Year
- Genre
- Synopsis

```{r urls}
html <- "https://github.com/addsding/data607/raw/main/assignment7/books.html"
xml <- "https://github.com/addsding/data607/raw/main/assignment7/books.xml"
json <- "https://github.com/addsding/data607/raw/main/assignment7/books.json"
```

Let's work on getting these into dataframes!

## Extracting the Data

### HTML

I'll be using the `rvest` package for this portion of the assignment.

```{r html}
html_df <- as.data.frame(read_html(html) |> html_table(fill=TRUE))
head(html_df)
```

Looks pretty good! One of the column names though has a period rather than a space/underscore -- let's change that! 

```{r clean_html}
html_df <- html_df |> rename("Publish_Year" = "Publish.Year")
head(html_df)
```

Seems relatively clean! For the `Author` column, we can try splitting that into 2 columns just so two authors aren't in one column for the third book.

```{r html_split_author}
html_df <- html_df |> 
  separate_wider_delim(Author, delim=", ", names = c("Author_1", "Author_2"), too_few = "align_start")

head(html_df)
```

Looks good! Let's move onto `json`.

### JSON

I'll be using the `jsonlite` package for this portion of the assignment.

```{r json}
json_df <- jsonlite::fromJSON(json)
json_df <- as.data.frame(json_df)
head(json_df)
```

For the `author` column, the third book causes it to be a list since there are two authors. Let's try to split that column into two to match the HTML data frame.

```{r json_split_author}
json_df <- json_df |> unnest_wider(books.author, names_sep="author")
head(json_df)
```

Looks perfect! Now just clean up the column names and it should be good to go.

```{r clean_json}
json_df <- json_df |> rename("title" = "books.title",
                             "author_1" = "books.authorauthor1",
                             "author_2" = "books.authorauthor2",
                             "publish_year" = "books.publishYear",
                             "genre" = "books.genre",
                             "synopsis" = "books.synopsis"
                             )
head(json_df)
```

Awesome, XML next.

### XML

I'll be using the `xml2` and `tibble` packages for this portion of the assignment.

```{r xml}
# get the XML
xml_books <- read_xml(xml)

# pull each column
title <- xml_text(xml_find_all(xml_books, xpath = "//title"))
author <- xml_text(xml_find_all(xml_books, xpath = "//author"))
year <- xml_text(xml_find_all(xml_books, xpath = "//publishYear"))
genre <- xml_text(xml_find_all(xml_books, xpath = "//genre"))
synopsis <- xml_text(xml_find_all(xml_books, xpath = "//synopsis"))

# turn it into a df; for author, it seems that we have to append the 2nd author for the third book separately
xml_df <- tibble(title = title
                 , author_1 = author[1:3]
                 , publish_year = year
                 , genre = genre
                 , synopsis = synopsis
                 )
author_2 <- c(NA, NA, author[4])
xml_df$author_2 <- author_2

head(xml_df)

```

Looks perfect!

## Conclusion

With these methods, I would say the data frames are basically the same due to my own manipulation. From the start, the data frames were certainly not exactly the same. For example, I do see now that the XML method does not make `publish_year` an int, however the data could be cleaned up a little more for sure to get things correct. The order of the columns is also not the same for XML due to how I structured it. I think ultimately, most fields would be imported into a similar data frame, it's just lists that make things complicated too as well as object types.

Working with each data structure, I found HTMl to be the easiest to work with due to the simplicity of the package and the fact that it kind of already starts as a table to begin with anyway. JSON and XML are similar in nature with how they're structured and how they're created, but the packages vary in difficulty to use overall. 

I think the code I've written can be reworked for multiple books with more fields as well for JSON and HTML, however the way I implemented the XML variation makes it a little more work if there are more columns to work with. It's also very much not repeatable if the books I had were in a different order -- I would have to rewrite it to fit my use case depending on the data presented.

Overall, I found this exercise to be interesting and fun in its own way as I got to work with different data structures and found a way to make the same three data frames!