---
title: "Assignment 3"
author: "Alice Ding"
date: "2023-02-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 
Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r majors}
majors <- read.csv("majors-list.csv")
to_find <- c("DATA", "STATISTICS")
matches <- unique(grep(paste(to_find, collapse="|"), majors$Major, value=TRUE))
matches
```

There are three majors that contain either "DATA" or "STATISTICS".

## Question 2
Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

> I'm going to assume that 

```{r orig}
fruits <- c("bell pepper","bilberry", "blackberry","blood orange","blueberry","cantaloupe", "chili pepper", "cloudberry",  "elderberry", "lime", "lychee" , "mulberry" , "olive", "salal berry")
fruits
```
```{r transform}
cat(paste0("c(", paste0(sep = '"', fruits, collapse = ', ', sep='"'),paste(")")))
```

## Question 3
Describe, in words, what these expressions will match:

- (.)\1\1
- "(.)(.)\\2\\1"
- (..)\1
- "(.).\\1.\\1"
- "(.)(.)(.).*\\3\\2\\1"

## Question 4

Construct regular expressions to match words that:

- Start and end with the same character.
- Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
- Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)